Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-04-04 src/tex/ms.tex
notable events done

Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, GW190521 and GW191105.
When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$; the second shows signs of potential noise anomalies; the third is representative of a class of heavy events with strongly bimodal $\kappa$ posteriors; and the fourth of a smaller class of lighter events with slightly bimodal posteriors.
Through these examples, we elucidate the interaction between $\kappa$ and the source luminosity distance, inclination and spin precession.

@@

% Case: GW191105
\subsubsection{GW191105}
\label{sec:GW191105}

The situation is similar for GW191105 (Fig.~\ref{fig:corner_GW191105}).

\begin{figure}
    \script{corner_GW191105.py}
    \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
    \caption{
        GW191105 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
        The posterior on $\chi_{eff}$ and $\kappa$ from our birefringence analysis is shown in the top right panel.
    }
    \label{fig:corner_GW191105}
\end{figure}

@@

\section{Extended corner plot for GW170818}
\label{sec:corner_GW170818_appendix}

@@

\begin{figure*}[h]
    \script{corner_GW170818_appendix.py}
    \includegraphics[width=\textwidth]{figures/corner_GW170818_appendix.pdf}
    \caption{
        Extended corner plot for GW170818: a supplement to Fig.~\ref{fig:corner_GW170818} discussed in Appendix~\ref{sec:corner_GW170818_appendix}.
    }
    \label{fig:corner_GW170818_appendix}
\end{figure*}

\section{Extended corner plot for GW190521}
\label{sec:corner_GW190521_appendix}

\begin{figure*}[h]
    \script{corner_GW190521_appendix.py}
    \includegraphics[width=\textwidth]{figures/corner_GW190521_appendix.pdf}
    \caption{
        Extended corner plot for GW190521: a supplement to Fig.~\ref{fig:corner_GW190521} discussed in Appendix~\ref{sec:corner_GW190521_appendix}.
    }
    \label{fig:corner_GW190521_appendix}
\end{figure*}

\section{Extended corner plot for GW191105}
\label{sec:corner_GW191105_appendix}

\begin{figure*}[h]
    \script{corner_GW191105_appendix.py}
    \includegraphics[width=\textwidth]{figures/corner_GW191105_appendix.pdf}
    \caption{
        Extended corner plot for GW191105: a supplement to Fig.~\ref{fig:corner_GW191105} discussed in Appendix~\ref{sec:corner_GW191105_appendix}.
    }
    \label{fig:corner_GW191105_appendix}
\end{figure*}

\section{Extended corner plot for GW200129}
\label{sec:corner_GW200129_appendix}

\begin{figure*}[h]
    \script{corner_GW200129_appendix.py}
    \includegraphics[width=\textwidth]{figures/corner_GW200129_appendix.pdf}
    \caption{
        Extended corner plot for GW200129: a supplement to Fig.~\ref{fig:corner_GW200129} discussed in Appendix~\ref{sec:corner_GW200129_appendix}.
    }
    \label{fig:corner_GW200129_appendix}
\end{figure*}

>>>>

Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW190521 and GW200129.
When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$; the third is representative of a class of events with bimodal $\kappa$ posteriors; and the third shows signs of potential noise anomalies.
Through these examples, we elucidate the interaction between $\kappa$ and the source luminosity distance, inclination and spin parameters.

@@

% % Case: GW191105
% \subsubsection{GW191105}
% \label{sec:GW191105}
% 
% The situation is similar for GW191105 (Fig.~\ref{fig:corner_GW191105}).
% 
% \begin{figure}
%     \script{corner_GW191105.py}
%     \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
%     \caption{
%         GW191105 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
%         The posterior on $\chi_{eff}$ and $\kappa$ from our birefringence analysis is shown in the top right panel.
%     }
%     \label{fig:corner_GW191105}
% \end{figure}

@@

\section{Extended results for notable events}

\subsection{GW170818}
\label{sec:corner_GW170818_appendix}

@@

\begin{figure*}[h]
    \script{corner_GW170818_appendix.py}
    \includegraphics[width=\textwidth]{figures/corner_GW170818_appendix.pdf}
    \caption{
        Extended corner plot for GW170818: a supplement to Fig.~\ref{fig:corner_GW170818} discussed in Appendix~\ref{sec:corner_GW170818_appendix}. The shaded regions contain 90\% and 39.35\% ($1\sigma$) of the probability mass.
        The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.
    }
    \label{fig:corner_GW170818_appendix}
\end{figure*}

\subsection{GW190521}
\label{sec:corner_GW190521_appendix}

As in the previous section, here we display additional parameters from the GW190521 measurement in Fig.~\ref{fig:corner_GW190521_appendix}, expanding upon Fig.~\ref{fig:corner_GW190521} in the main text.
The reference \ac{GR} analysis, obtained with the \textsc{IMRPhenomXPHM} waveform, favors near extremal spins, in particular for the heavier component, and rules out nonspinning objects (i.e., $\chi_1 = \chi_2 = 0$) with $\geq 90\%$ credibility.
This can be seen most clearly in the joint posterior for $\chi_1$ and $\chi_2$, rather than in the respective 1D marginals because the \acp{BH} in this system were inferred to have equal masses \cite{Biscoveanu:2020are}.

Similar to GW170818, the \ac{GR} analysis for GW190521 also provides an informative measurement of the polarization angle $\psi$ and reference phase $\phi_{\rm ref}$.
However, the fact that $\psi$ is constrained does not alone imply that the degeneracy between right- and left-handed states is broken (for a nonprecessing source, $\psi$ is related to the difference in phase between the circular polarization modes \cite{Isi:2022mbx}).
Additionally, the posterior for tilt angle of the primary \ac{BH}, $\theta_1$, shows support for an antialigned spin ($\cos\theta_1 \approx -1$).
For this events, spin and phase angles are referred to 10 Hz at the detector.

Unlike the \ac{GR} analysis, the birefringent analysis does not favor extremal spins and is, in fact, consistent with nonspinning objects within 90\% credibility (lighter blue region encloses $\chi_1 = \chi_2 = 0$ in the respective panel of Fig.~\ref{fig:corner_GW190521_appendix}).
In Fig.~\ref{fig:corner_GW190521}, this same fact manifested as a $\chi_{\rm eff}$ posterior with higher support for the origin.
As we discussed in Sec.~\ref{sec:GW190521}, this suggests that some of the spin effects in these data are being absorbed by a nonzero $\kappa$.
Other than the spins, the birefringent analysis differs most notably from the \ac{GR} one in the inference of the luminosity distance, which is favored to be much greater.
See Sec.~\ref{sec:GW190521} in the main text for related exposition.

\begin{figure*}[h]
    \script{corner_GW190521_appendix.py}
    \includegraphics[width=\textwidth]{figures/corner_GW190521_appendix.pdf}
    \caption{
        Extended corner plot for GW190521: a supplement to Fig.~\ref{fig:corner_GW190521} discussed in Appendix~\ref{sec:corner_GW190521_appendix}. The shaded regions contain 90\% and 39.35\% ($1\sigma$) of
 the probability mass.
        The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.
    }
    \label{fig:corner_GW190521_appendix}
\end{figure*}

% \subsection{Extended corner plot for GW191105}
% \label{sec:corner_GW191105_appendix}
% 
% \begin{figure*}[h]
%     \script{corner_GW191105_appendix.py}
%     \includegraphics[width=\textwidth]{figures/corner_GW191105_appendix.pdf}
%     \caption{
%         Extended corner plot for GW191105: a supplement to Fig.~\ref{fig:corner_GW191105} discussed in Appendix~\ref{sec:corner_GW191105_appendix}.
%     }
%     \label{fig:corner_GW191105_appendix}
% \end{figure*}

% \section{Extended corner plot for GW200129}
% \label{sec:corner_GW200129_appendix}
% 
% \begin{figure*}[h]
%     \script{corner_GW200129_appendix.py}
%     \includegraphics[width=\textwidth]{figures/corner_GW200129_appendix.pdf}
%     \caption{
%         Extended corner plot for GW200129: a supplement to Fig.~\ref{fig:corner_GW200129} discussed in Appendix~\ref{sec:corner_GW200129_appendix}.
%     }
%     \label{fig:corner_GW200129_appendix}
% \end{figure*}
